---
title: "HW1"
author: "Chunjie Nan"
date: "6/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


2.1 
Use the help function to explore what the series gold, woolyrnq and gas represent.

a. Use autoplot() to plot each of these in separate plots.

b. What is the frequency of each series? Hint: apply the frequency() function.

c. Use which.max() to spot the outlier in the gold series. Which observation was it?


2.3 
Download some monthly Australian retail data from the book website. These represent retail sales in various categories for different Australian states, and are stored in a MS-Excel file.

a. You can read the data into R with the following script:
```{r}
retaildata <- readxl::read_excel("retail.xlsx", skip=1)
```
The second argument (skip=1) is required because the Excel sheet has two header rows.


b. Select one of the time series as follows (but replace the column name with your own chosen column):
```{r}
myts <- ts(retaildata[,"A3349873A"],
  frequency=12, start=c(1982,4))
```

c. Explore your chosen retail time series using the following functions:
autoplot(), ggseasonplot(), ggsubseriesplot(), gglagplot(), ggAcf()

Can you spot any seasonality, cyclicity and trend? 

What do you learn about the series?


6.2
The plastics data set consists of the monthly sales (in thousands) of product A for a plastics manufacturer for five years.

a. Plot the time series of sales of product A. Can you identify seasonal fluctuations and/or a trend-cycle?

b. Use a classical multiplicative decomposition to calculate the trend-cycle and seasonal indices.

c. Do the results support the graphical interpretation from part a?

d. Compute and plot the seasonally adjusted data.

e. Change one observation to be an outlier (e.g., add 500 to one observation), and recompute the seasonally adjusted data. What is the effect of the outlier?

f. Does it make any difference if the outlier is near the end rather than in the middle of the time series?

